---
title: "Data 607 Homework 5"
author: "David Blumenstiel"
date: "2/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Tidying Data
Here, we will take some messy data and tidy it

#### Importing the untidy data
```{r}
#Imports the csv (which I put on github)
dirty <- read.csv('https://raw.githubusercontent.com/davidblumenstiel/data/master/Flight.csv')
dirty
```

#### Cleaning up
```{r}
#Removing blank rows
df <- na.omit(dirty)
df
#Adding column names

names(df)[1:2] <- c("Airline", "Status")

#Appending missing values in "Airline"
#This is a little overkill, but would work if the dataset were larger
j = 0
for (i in df$Airline) {
  j = j + 1
  if (i != "") {
    x = i
  }
  else {
    df$Airline[j] <- x
  }
}
df
#Tidying up the destinations
clean <- gather(df,"Destination","Number Flights",3:7)
head(clean)
```

#### Comparing arrivals:

##### What proportion of flights were on time/delayed for each destination across each airline?
```{r}
#Creates a vector of the total number of flights per airline and destination
totals <- rep(clean$`Number Flights`[clean$Status == "on time"] + clean$`Number Flights`[clean$Status == "delayed"], each = 2)

#Initiates new column (Proportion Ontime/Delayed)
clean["Proportion Ontime/Delayed"] <- NA

#Fills the new column
clean$`Proportion Ontime/Delayed` <- clean$`Number Flights`/totals
head(clean)
```


##### A little visualization

```{r}
delayed <- filter(clean,Status == "delayed")

ggplot(delayed, aes(Airline, `Proportion Ontime/Delayed`)) + geom_boxplot()
```

Although there isn't too much data to work with, it appears that the 'AMWEST' airline had a higher proportion of delayed flights than did the 'ALASKA' airline.