---
title: "Data 607 Project 1"
author: "David Blumenstiel"
date: "2/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
### Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
### For the first player, the information would be:
### Gary Hua, ON, 6.0, 1794, 1605







### Importing the text file from GitHub
```{r}
#I posted the data to GitHub; this loads it (without spaces or the first row of dots)
df <- read.csv("https://raw.githubusercontent.com/davidblumenstiel/data/master/tournamentinfo.txt", header = F, skip = 1, strip.white = T)
head(df)

```

### Removing the junk we don't need and concatenating every other row
```{r}
#Removing the dotted lines (which occur every third line)
df1 <- data.frame(df[-seq(0,nrow(df),3), ], "")



#Concatenating every other row
df2 <- data.frame(cbind(df1[ (seq(nrow(df1)) %% 2)==1,], df1[ (seq(nrow(df1)) %% 2)==0,]))
df2$X...1 <- df2$X.. <- NULL
df2 <- df2$all <- paste(df2$df..seq.0..nrow.df...3...., df2$df..seq.0..nrow.df...3.....1)


#Delimiting by '|' into the data frame: players
players <- data.frame(do.call('rbind', strsplit(as.character(df2),'|',fixed=TRUE)))


head(players)
```

### Subsetting the data and renaming the columns
```{r}
#Getting rid of columns we aren't interested in
players <- players[ ,-c(13:20)]

#Renaming the columns and getting rid of the first row
colnames(players) <- c("Player_ID","Player_Name", "Total_Score",  "Round_1_Opponent","Round_2_Opponent","Round_3_Opponent","Round_4_Opponent","Round_5_Opponent","Round_6_Opponent","Round_7_Opponent","State","PreRating")
players <- players[-1,]
head(players)
```

### Changing opponent columns to opponent IDs
```{r}
#Removes the Win/Loss info and coerces remainder to an int
players$Round_1_Opponent <- as.integer(gsub(". ","",players$Round_1_Opponent))
players$Round_2_Opponent <- as.integer(gsub(". ","",players$Round_2_Opponent))
players$Round_3_Opponent <- as.integer(gsub(". ","",players$Round_3_Opponent))
players$Round_4_Opponent <- as.integer(gsub(". ","",players$Round_4_Opponent))
players$Round_5_Opponent <- as.integer(gsub(". ","",players$Round_5_Opponent))
players$Round_6_Opponent <- as.integer(gsub(". ","",players$Round_6_Opponent))
players$Round_7_Opponent <- as.integer(gsub(". ","",players$Round_7_Opponent))
head(players)
```


### Extracting the Pre-Ratings from the 'PreRating' column, and changing the data type to numeric
```{r}
library(stringr)
#Filtering out everything but the pre-rating in it's respective column
players$PreRating <- gsub(" ","",players$PreRating)
players$PreRating <- gsub("-.*","",players$PreRating)
players$PreRating <- gsub(".*:","",players$PreRating)
players$PreRating <- gsub("P.*","",players$PreRating)

#Changing the type to numeric
players$PreRating <- as.numeric(players$PreRating)

head(players)
```

### Making a new column of average opponents' pre-ratings (Mean_Opponent_PreRatings)
```{r}
library(gbutils)
#Changing Player_ID to integer
players$Player_ID <- as.integer(as.character(players$Player_ID))

#Initializing an empty column
players$Mean_Opponent_PreRatings <- NA

#Loops through the dataset, calculate's average pre-ratings for all players except the one in question, and appends it to the new column
for (i in 1:nrow(players)) {
  #Initializes an empty vector
  opts = c()
  #Loops through columns within a player's row to build a vector of their opponent's mean ratings
  for (k in c(4:10)) {
    if (isNA(players[i,k])== FALSE) {
      opts = c(opts, players$PreRating[players$Player_ID==players[i,k]])
    }
  }
  #Appends the mean of a player's opponent's pre-ratings to the respective column
  players[i,"Mean_Opponent_PreRatings"] <- round(mean(opts), digits = 0)

}

head(players)
```

### Cleaning up: removing the opponent columns and rearranging
```{r}
#Removing unnecessary columns
players[,c(1,4:10)] <- NULL

#Reordering columns
players <- players[,c(1,3,2,4,5)]

row.names(players) <- NULL
head(players)
```

### Writing to a csv
```{r}
#Creates a csv file with the desired info in the workfolder
write.csv(players, file = "Chess_Tournament_Player_Info.csv")
```



