---
title: "Data 607 Project 2"
author: "David Blumenstiel"
date: "3/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
```

# Choose 3 wide data sets, tidy, and analyze


## Dataset #1: Agricultural Land Values
### This dataset contains the values of agricultural land from 2015-2019 across different states and regions


#### Importing from csv and basic cleaning
```{r}
#Importing
agland <- read.csv("https://raw.githubusercontent.com/davidblumenstiel/data/master/Agricultural%20Land%20Values.csv")
head(agland)
#Renaming columns
names(agland)[1:7] <- c("Region/State","2015","2016","2017","2018","2019","%Change2018-2019")

#Getting rid of blank rows
agland <- na.omit(agland)

#Reformatting columns to their appropriate types
#Need to remove the '.....'s from Column 1
agland[,1] <- gsub("\\.*:","",as.character(agland[,1]))

#Changing the value by year columns to numeric; also removing the  ','s
agland[,2] <- as.numeric(gsub(",","",as.character(agland[,2])))
agland[,3] <- as.numeric(gsub(",","",as.character(agland[,3])))
agland[,4] <- as.numeric(gsub(",","",as.character(agland[,4])))
agland[,5] <- as.numeric(gsub(",","",as.character(agland[,5])))
agland[,6] <- as.numeric(gsub(",","",as.character(agland[,6])))
head(agland)
```
The year columns would make an easy target for tidying.


#### Tidying
##### Making columns for the dates and the land values
```{r}
agland <- gather(agland,"Year","Value",2:6)
head(agland)
```


#### Perform the requested analysis
##### No analysis was requested, so instead we will find the median land value per state across all 5 years

```{r}
#Easy to do with dplyr
medAgland <- agland %>%
  group_by(`Region/State`) %>%
  summarize(Median_Value = median(Value))

head(medAgland)


#Now what we have the median values, we can use ggplot2 to make a nice plot
ggplot(medAgland, aes(x = reorder(`Region/State`,Median_Value), y = Median_Value, fill = `Region/State`)) +
  coord_flip() +
  geom_col() +
  ggtitle("Median Land Value Across All Years")+
  xlab("Median Value") +
  ylab("State/Region") +
  theme(legend.position = "none")
 

```

Rhode Island has the highest median land value ($14900), and North Dakota has the lowest (1740)






##### How do the land values change over the years
```{r}
#Good time to use box plots
ggplot(agland,aes(x = Year, y = Value, fill = Year)) +
  geom_boxplot() +
  scale_fill_brewer(palette=1)+ 
  ggtitle("Land Value by Year")
  
```

Each of the years are pretty similar in terms of land value, with a slight increase per year.  One interesting observation is that the IQR is tightening as time goes on.



## Dataset #2: Student Performance
### This is a dataset containing student test scores on two different tests, across two different years
### Disclaimer: I did not make the csv for this; it was alerady on github


#### Importing the dataset
```{r}
#Importing
stuper <- read.csv("https://gist.githubusercontent.com/Kimmirikwa/b69d0ea134820ea52f8481991ffae93e/raw/4db7b1698035ee29885d10e1a59bd902716ae168/student_results.csv")
stuper


```
This dataset arrives much cleaner than the last one, however the term columns could be combined, and the sex and age column can be seperated into 2


#### Spliting and tidying columns
```{r}
#Spliting the sex.and.age column with dplyr: separate
stuper <- stuper %>%
  separate(sex.and.age,c("sex","age"),"_")
stuper

#Combining the term columns with tidyr: gather
stuper <- gather(stuper,"term","score",7:9)
head(stuper)

```

#### Analysis
##### Determining student performance changes during later terms
```{r}
#Can make a new output dataframe for clarification:
#One row per student per test
stuimp <- data.frame(matrix(data = NA,nrow = 10,ncol=3))
names(stuimp) <- c("Student","Test Number","Ratio of Improvment Between Terms 1 and 3")
#Takes student names from the tidy dataset
stuimp$Student <- as.character(stuper$name[1:10])

#Makes a vector of the percent change ia student's score for test 1, and appends it to it's new column int he output dataframe
testImprove <- (stuper$score[stuper$term == "term.3" & stuper$test.number == "test 1"] -
  stuper$score[stuper$term == "term.1" & stuper$test.number == "test 1"]) / 
  stuper$score[stuper$term == "term.3" & stuper$test.number == "test 1"]
  
#Does the same thing for test 2 and adds it to the vector above
testImprove <- c(testImprove,(stuper$score[stuper$term == "term.3" & stuper$test.number == "test 2"] -
  stuper$score[stuper$term == "term.1" & stuper$test.number == "test 2"]) / 
  stuper$score[stuper$term == "term.3" & stuper$test.number == "test 2"])

#Appends testImprove to the output dataframe
stuimp$`Ratio of Improvment Between Terms 1 and 3` <- testImprove

#Adds the test number
stuimp$`Test Number` <- c(rep(1,5),rep(2,5))

stuimp
```
The output is a small, tidy table of only the student changes in performance

##### We can also plot the performance changes by student for each test
```{r}
#Lets also determine the median performance change for all students on each test
medperform1 <- median(stuimp$`Ratio of Improvment Between Terms 1 and 3`[stuimp$`Test Number` == 1])
medperform2 <- median(stuimp$`Ratio of Improvment Between Terms 1 and 3`[stuimp$`Test Number` == 2])

#Test 1
barplot(100 * c(stuimp$`Ratio of Improvment Between Terms 1 and 3`[stuimp$`Test Number` == 1], medperform1),names =
          c(stuimp$Student[stuimp$`Test Number` == 1],"Median"),xlab = "Student", ylab = "% Performance Change",main = "Test 1 Improvements",col=rgb(0.5,0.2,0.1,0.9))

#Test 2
barplot(100 * c(stuimp$`Ratio of Improvment Between Terms 1 and 3`[stuimp$`Test Number` == 2], medperform2),names =
          c(stuimp$Student[stuimp$`Test Number` == 2],"Median"),xlab = "Student", ylab = "% Performance Change",main = "Test 2 Improvements",col=rgb(0.1,0.0,0.6,0.2))

```

Overall, students' median performance increased on both tests, but more on test 2.  Linda and mary were the only ones who performed signifcantly worse on the later tests (1 and 2 respectively).



## Dataset #3: Bob Ross
### This is a dataset of Bob Ross TV Episodes, the painting in the episode, and all the features in the painting
### Disclaimer: I did not make the csv for this; I did upload it to github however


#### Importing the dataset
```{r}
bobross <-read.csv("https://raw.githubusercontent.com/davidblumenstiel/data/master/bobross.csv")
head(bobross)
```
This has way to many columns; luckily they all have the same kind of observations, and would be easy to tidy.

#### Combining all the 'features' into one column
```{r}
#Going to use pivot_gather here; pretty much the same thing

bobross <- bobross %>%
  pivot_longer(-c("EPISODE","TITLE"), names_to = "Feature", values_to = "Included")

head(bobross)
```
Much easier to read now, although our 400 observations have turned into 27000.


#### Now that it's easy to use our super tidy (but long) database for analysis


##### What are the most used features
```{r fig.height=20}
#Credit where credit is due: Leo Yi assisted greatly with this chunk
group_by(bobross,Feature) %>%
  summarize(Occurances = sum(Included)) %>%
  ggplot(aes(x = Feature, y = Occurances, fill = Feature)) +
  geom_col() +
  theme_linedraw() +
  theme(legend.position = "none",axis.text=element_text(size=10),
        axis.title=element_text(size=20))  +
  coord_flip() +
  
  ggtitle("Number of Features Uses Across All Episodes")+
  theme(plot.title = element_text(size=15)) 
 
  

```

It appears that tree(s) are the most common features in his paintings, followed by clouds if one discounts the seperate types of trees.


##### What episodes include fire, and what are their features?
```{r fig.height=10}
fire <- bobross %>%
  filter(Feature == 'FIRE') %>%
  filter(Included == 1)

fire[,1:2]

```
##### It turns out only the epsidoe S03E10, "CAMPFIRE" includes fire, but what other features does it have?
```{r}
bobross %>%
  filter(EPISODE == "S03E10") %>%
  filter(Included == 1)
```
We can probably say that the painting in this episode includes a person outdoors by the fire with a lake nearby, likely in some wooded setting.

